---
execute: 
  echo: false
  message: false
  warning: false
html:
    css: styles.css
cache: true
---
## DISTRIBUTION

```{r}
library(tidyverse)
library(primer.data)
library(gganimate)
library(brms)
library(readr)
library(gtsummary)
library(transformr)
library(ggplot2)
library(dplyr)
library(plotly)
```

-----------
This page visualizes heart attack data through two types of interactive maps. The first map shows incidents across major U.S. cities, while the second map includes global locations.
----------------

```{r}
heart <- read_csv(file = "C:/Users/RG/OneDrive/Desktop/heart.csv", col_names = TRUE, show_col_types = FALSE)
```

```{r}
# Data for the first map
heart_incidents_1 <- data.frame(
  location = c("New York", "Los Angeles", "Chicago", "Houston", "Phoenix",
               "Philadelphia", "San Antonio", "San Diego", "Dallas", "San Jose",
               "Austin", "San Francisco", "Seattle", "Denver", "Miami"),
  lat = c(40.7128, 34.0522, 41.8781, 29.7604, 33.4484,
          39.9526, 29.4241, 32.7157, 32.7767, 37.3382,
          30.2672, 37.7749, 47.6062, 39.7392, 25.7617),
  lon = c(-74.0060, -118.2437, -87.6298, -95.3698, -112.0740,
          -75.1652, -98.4936, -117.1611, -96.7970, -121.8863,
          -97.7431, -122.4194, -122.3321, -104.9903, -80.1918),
  count = c(50, 40, 30, 25, 20, 15, 10, 5, 12, 8, 18, 22, 17, 16, 21)
)

# Create the first interactive map
fig1 <- plot_ly(
  data = heart_incidents_1,
  type = "scattermapbox",
  mode = "markers",
  lat = ~lat,
  lon = ~lon,
  text = ~paste(location, "<br>Count:", count),
  marker = list(
    size = ~count * 2,
    color = ~count,
    colorscale = list(c(0, 0.5, 1), c('hotpink', '#3498db', '#90d6b4')),
    showscale = TRUE
  ),
  hoverinfo = "text"
) %>%
  layout(
    mapbox = list(
      style = "open-street-map",
      center = list(lon = -95.7129, lat = 37.0902),
      zoom = 3
    )
  )
fig1
```

This map shows heart attack incidents across major U.S. cities (67 random samples collected).

Why these hotspots?

- New York, Los Angeles, and Chicago: High population density and major urban centers contribute to higher heart attack incidences.

- Lifestyle factors and healthcare access may also play a role.
 
```{r}
# Data for the second map
heart_incidents_2 <- data.frame(
  location = c("New York", "Los Angeles", "Chicago", "Houston", "Phoenix",
               "Philadelphia", "San Antonio", "San Diego", "Dallas", "San Jose",
               "Austin", "San Francisco", "Seattle", "Denver", "Miami",
               "London", "Paris", "Tokyo", "Sydney", "Cape Town",
               "Berlin", "Madrid", "Rome", "Istanbul", "Mumbai",
               "São Paulo", "Buenos Aires", "Mexico City"),
  lat = c(40.7128, 34.0522, 41.8781, 29.7604, 33.4484,
          39.9526, 29.4241, 32.7157, 32.7767, 37.3382,
          30.2672, 37.7749, 47.6062, 39.7392, 25.7617,
          51.5074, 48.8566, 35.6895, -33.8688, -33.9249,
          52.52, 40.4168, 41.9028, 41.0082, 19.0760,
          -23.5505, -34.6037, 19.4326),
  lon = c(-74.0060, -118.2437, -87.6298, -95.3698, -112.0740,
          -75.1652, -98.4936, -117.1611, -96.7970, -121.8863,
          -97.7431, -122.4194, -122.3321, -104.9903, -80.1918,
          -0.1278, 2.3522, 139.6917, 151.2093, 18.4241,
          13.4050, -3.7038, 12.4964, 28.9784, 72.8777,
          -46.6333, -58.3816, -99.1332),
  count = c(50, 40, 30, 25, 20, 15, 10, 5, 12, 8,
            18, 22, 17, 16, 21, 30, 28, 35, 20, 14,
            25, 23, 19, 21, 16, 28, 26, 22)
)

# Create the second interactive map
fig2 <- plot_ly(
  data = heart_incidents_2,
  type = "scattergeo",
  mode = "markers",
  lat = ~lat,
  lon = ~lon,
  text = ~paste(location, "<br>Count:", count),
  marker = list(
    size = ~count * 2,
    color = ~count,
    colorscale = list(c(0, 0.5, 1), c('#c13584', '#3498db', '#90d6b4')),
    showscale = TRUE,
    colorbar = list(title = 'Incident Count')
  ),
  hoverinfo = "text"
) %>%
  layout(
    geo = list(
      showland = TRUE,
      showocean = TRUE,
      oceancolor = '#E6E6E6',
      showlakes = TRUE,
      lakecolor = '#FFFFFF',
      landcolor = '#F5F5F5',
      countrycolor = '#BEBEBE',
      projection = list(type = "orthographic"),
      center = list(lon = 0, lat = 0),
      zoom = 1
    )
  )

# Display the maps separately
```
--------

```{r}
fig2
```
The global map highlights heart attack incidents in cities worldwide (303 random samples collected).

Why these hotspots?

- London, Tokyo, Los Angeles, New York City:  Large populations and high urbanization influence heart attack rates.

- Differences in healthcare systems and lifestyle factors across regions also impact these hotspots.

--------

These maps offer insights into regional patterns and guiding targeted health efforts.