---
execute: 
  echo: false
  message: false
  warning: false
html:
    css: styles.css
cache: true
---


```{r}
library(tidyverse)
library(primer.data)
library(gganimate)
library(brms)
library(readr)
library(gtsummary)
library(transformr)
library(ggplot2)
library(dplyr)
library(plotly)
library(shiny)
```

```{r}
heart <- read_csv(file = "C:/Users/RG/OneDrive/Desktop/heart.csv", col_names = TRUE, show_col_types = FALSE)
```
## INTRODUCTION
To better understand heart attack risk, we look at how age, cholesterol, chest pain types, and heart rate affect the chance of having a heart attack. This analysis uses data from Kaggle. We aim to find these key predictors and see how they interact. Our goal is to answer the following question:

***What are the most significant predictors of heart attacks based on this dataset, and how do these factors interact with each other?***

Answering this question helps  healthcare professionals identify key risk factors and interactions for improved prevention and prediction.

---------

Gender is a significant factor in heart attack risk. This graph explores the frequency of heart attacks among males and females to identify patterns and disparities.
```{r}
ggplot(heart, aes(x = as.factor(sex), fill = as.factor(sex))) +
  geom_bar() +
  labs(title = "Frequency of Heart Attacks by Gender",
       x = "Gender",
       y = "Count",
       caption = "Data Source: Kaggle") +
  scale_fill_manual(values = c("0" = "hotpink", "1" = "cornflowerblue"),
                    labels = c("0" = "Female", "1" = "Male"),
                    name = "Gender") +
  theme_minimal()
```

- Males have more than double the frequency of heart attacks compared to females.
- Highlights gender disparities in heart attack prevalence.


----------------


This next animated plot explores the distribution of maximum heart rate achieved across different age groups, revealing important heart health patterns.
```{r}
heart <- heart %>%
  mutate(age_group = cut(age, breaks = c(0, 40, 50, 60, 70, 80, 100), 
                         labels = c("0-40", "40-50", "50-60", "60-70", "70-80", "80+")))

heart_summary <- heart %>%
  group_by(age_group) %>%
  summarize(avg_thalachh = mean(thalachh, na.rm = TRUE))

animated_plot <- ggplot(heart_summary, aes(x = age_group, y = avg_thalachh, fill = avg_thalachh)) +
  geom_col(show.legend = FALSE) +
  scale_fill_gradient(low = "lightpink", high = "cornflowerblue") + 
  labs(title = "Average Maximum Heart Rate Achieved by Age Group",
       x = "Age Group",
       y = "Average Maximum Heart Rate",
       fill = "Average Heart Rate") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  transition_states(age_group, transition_length = 2, state_length = 1) +
  enter_fade() +
  exit_fade() +
  labs(subtitle = "Age Group: {closest_state}")

animated_plot
```
- The plot shows that younger groups (0-40) have higher maximum heart rates, while older groups (60-70, 70-80) have lower rates.
- Important for assessing heart health across different age groups.

---------

This graph reveals the distribution of cholesterol levels among heart attack victims in the dataset.
```{r}
ggplot(heart, aes(x = chol)) +
  geom_histogram(binwidth = 20, fill = "#90d6b4") +
  labs(title = "Distribution of Cholesterol Levels",
       x = "Cholesterol",
       y = "Number of Patients", caption = "Data Source: Kaggle") +
  theme_minimal()
```

- Cholesterol levels are predominantly between 200 and 240 mg/dL.
- Indicates common cholesterol levels in the dataset.
- Guides general health recommendations and interventions.



[Learn more about Cholesterol and Heart Health](https://www.heart.org/en/health-topics/cholesterol/about-cholesterol#:~:text=Cholesterol%20circulates%20in%20the%20blood,you%20can%20know%20your%20levels.)

--------

